---
title: "Coursera_PracticalMachineLeaning"
author: "Anastasia Steshina"
date: "10/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Info

The main goal of the Project is to quantify how well an individual perform for a particular activity. This will be accomplished by training a prediction model on the accelerometer data using a random forest classifier.

The training data and the test data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

### Loading libraries, setting seed for reproducibility

```{r}
library(caret)
library(corrplot)
library(randomForest)
```

Loading data

```{r load, echo=FALSE }
train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",sep = ",", na.strings = c("", "NA"))
validation <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",sep = ",", na.strings = c("", "NA"))
```

Cleaning data:
```{r clean, cache=TRUE}
train_noNA <- train[ , colSums(is.na(train)) == 0] # Removing variables with NA
train_clean <- train_noNA[,-c(1:7)] # Removing unnessary columns

validation_noNA <- validation[ , colSums(is.na(validation)) == 0] # Removing variables with NA
validation_clean <- validation_noNA[,-c(1:7)] # Removing unnessary columns
```


Creating data partitions 70% training, 30% test
```{r partition, cache=TRUE}
set.seed(7)
inTrain = createDataPartition(y=train_clean$classe, p=0.7, list=FALSE)
training = train_clean[inTrain,]
testing = train_clean[-inTrain,]
```


Identifying variables that highly correlate with each other and can be removed before training the model
```{r correlation, cache=TRUE}
classeIndex <- which(names(training) == "classe")
correlationMatrix <- cor(training[, -classeIndex])
excludeColumns <- findCorrelation(correlationMatrix, cutoff=0.9, exact=TRUE) # Exclude highly correlated features
corrplot(correlationMatrix, method="color", type="lower", order="hclust", tl.cex=0.70, tl.col="black", tl.srt = 45, diag = FALSE)
training <- training[, -excludeColumns]
testing <- testing[, -excludeColumns]
```

Build prediction model using Random Forest:
```{r train, , cache=TRUE}
model <- randomForest(classe ~ .,   data=training, ntree = 50)
```


Evaluating the model:
```{r evaluate_model, , cache=TRUE}
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

The model is `r modelAcc` accurate on the testing set.

Predictions on the validation set:
```{r validation, , cache=TRUE}
predictions <- predict(model, validation)
predictions
```


